---
title: "P8105 HW1"
author: Shihui Zhu
date: 2021-09-21
output: github_document
---

# Problem 1

## Create a Data Frame
* a random sample of size 10 from a standard Normal distribution
```{r randome sample}
set.seed(1)
samp <- rnorm(10)
```

* a logical vector indicating whether elements of the sample are greater than 0
```{r logical vector}
log_vec <- samp > 0
```

* a character vector of length 10
```{r character vector}
char_vec <- c("This", "is", "a", "data", "science", "class", "please", "bring", 
            "your", "computer")
```

* a factor vector of length 10, with 3 different factor “levels”
```{r factor vector}
factor_vec <- factor(
  c('c_1','c_1','c_2','c_3','c_3','c_2','c_3','c_1','c_3','c_2'),
  levels = c('c_1','c_2','c_3')
  )
```

Create the dataframe
```{r dataframe}
df <- data.frame(samp, log_vec, char_vec, factor_vec)
df
```

## Take mean of each variable
```{r packages, results ='hide', message = FALSE}
# Load package
library(tidyverse)
```

```{r mean, collapse=TRUE}
# Mean of random sample
mean(dplyr::pull(df, 1))
# Mean of logical vector
mean(dplyr::pull(df, 2))
# Mean of char vector
mean(dplyr::pull(df, 3))
# Mean of factor vector
mean(dplyr::pull(df, 4))
```

The calculation for the mean of the first (random sample of 10 from a standard normal distribution) and second variable (logical vector) works. For 
character vector and factor vector, it does not work. 

## Convert Variables
```{r convert, eval=FALSE}
as.numeric(samp) # random vector to numeric
as.numeric(log_vec) # logical vector to numeric
as.numeric(char_vec) # character vector to numeric
as.numeric(factor_vec) # factor vector to numeric
```

There is a warning message showing for the character vector, saying NAs 
introduced by coercion. 

The `as.numeric()` function converts whatever data type to double, therefore:

* The random sample contains only the numeric values, so it does not change

* Logical vector contains flags of `TRUE` and `FALSE`, therefore it converts to vector of 1's and 0's

* Character vector contains string of characters, which do not have numeric meaning, therefore such coercion makes R assign null values to them, i.e. `as.numeric()` does not work

* For factor vector, `as.numeric()` return the underlying numeric representation, which is often meaningless as it may not correspond to the factor levels. 

This explains the NA's I got when I tried to take the mean of character vector and factor vector, since `mean()` only takes numeric values. 

# Problem 2
## Load Dataset
```{r load_data, message=FALSE}
data("penguins", package = "palmerpenguins")
```
## Description

The penguins dataset contains 8 different categories, `r names(penguins)`. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns. The data are collected from year `r unique(dplyr::pull(penguins, "year"))`. There are `r length(levels(dplyr::pull(penguins, "species")))` species of penguins in total, from `r length(levels(dplyr::pull(penguins, "island")))` different islands. The mean flipper length of the penguins is `r mean(dplyr::pull(penguins, "flipper_length_mm"), na.rm = TRUE)` mm, with a standard deviation of `r sd(dplyr::pull(penguins, "flipper_length_mm"), na.rm = TRUE)` mm; the mean bill length of the penguins is `r mean(dplyr::pull(penguins, "bill_length_mm"), na.rm = TRUE)` mm, with a standard deviation of `r sd(dplyr::pull(penguins, "bill_length_mm"), na.rm = TRUE)` mm; the mean bill depth of the penguins is `r mean(dplyr::pull(penguins, "bill_depth_mm"), na.rm = TRUE)` mm, with a standard deviation of `r sd(dplyr::pull(penguins, "bill_depth_mm"), na.rm = TRUE)` mm; and the mean body mass is `r mean(dplyr::pull(penguins, "body_mass_g"), na.rm = TRUE)` g, with a standard deviation of `r sd(dplyr::pull(penguins,"body_mass_g"), na.rm = TRUE)` g.

## Plots
The next chunk plots the scatter plot of flipper length v.s. bill length, colored according to species. 
```{r plot}
ggplot(penguins, 
       aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() + xlab("Bill Length (mm)") + ylab("Flipper Length (mm)") +
  ggtitle("Scatter Plot of Flipper Length v.s. Bill Length")
```

Finally, save the scatter plot:
```{r save, warning=FALSE}
ggsave("hw1_p2_scatter_plot.jpg")
```
