---
title: "P8105 HW1"
author: Shihui Zhu
date: 2021-09-21
output: github_document
---

# Problem 1

## Create a Data Frame
* a random sample of size 10 from a standard Normal distribution
```{r randome sample}
set.seed(1)
samp <- rnorm(10)
```

* a logical vector indicating whether elements of the sample are greater than 0
```{r logical vector}
log_vec <- samp > 0
```

* a character vector of length 10
```{r character vector}
char_vec <- c("This", "is", "a", "data", "science", "class", "please", "bring", 
            "your", "computer")
```

* a factor vector of length 10, with 3 different factor “levels”
```{r factor vector}
factor_vec <- factor(c(1,1,2,3,3,2,3,1,3,2))
```


```{r dataframe}
df <- data.frame(samp, log_vec, char_vec, factor_vec)
df
```

## Take mean of each variable
```{r packages, results ='hide', message = FALSE}
library(tidyverse)
```

```{r mean, collapse=TRUE}
# Mean of random sample
mean(dplyr::pull(df, 1))
# Mean of logical vector
mean(dplyr::pull(df, 2))
# Mean of char vector
mean(dplyr::pull(df, 3))
# Mean of factor vector
mean(dplyr::pull(df, 4))
```

The calculation for the mean of the first (randome sample of 10 from a standard normal distribution) and second variable (logical vector) works. For 
character vector and factor vector, it does not work. 

## Convert Variables
```{r convert, results='hide', warning=FALSE}
as.numeric(log_vec)
as.numeric(char_vec)
as.numeric(factor_vec)
```

There is a warning message showing for the character vector, saying NAs 
introduced by coercion. The characters do not have meanings in numeric values, 
therefore they cannot be converted to numbers, and such coercion makes R 
assign null values to them. This explains the null value I got when I tried to 
take the mean of it. 

# Problem 2
```{r load_data}
data("penguins", package = "palmerpenguins")
```
## Description
* Summary of the dataset
```{r describe}
# summary of the dataset
summary(penguins)

```
* Variable names
```{r}
names(penguins)
```
* Number of rows
```{r}
nrow(penguins)
```
* Number of cols
```{r}
ncol(penguins)
```
